---
title: "vfrench3_OriginalHomeworkCode_02"
author: "VIctoria French"
date: "9/22/2021"
output: html_document
---

# Homework 2: Titis in the Mist 

Every Saturday morning, at the same time, a primatologist goes and sits in the forest to listen for titi monkey calls, counting the number of calls they hear in a 2 hour window from 5am to 7am. Based on previous knowledge, she believes that the mean number of calls she will hear in that time is exactly 15. Let X represent the appropriate Poisson random variable of the number of calls heard in each monitoring session.

## Notes on the Poisson Distribution 

- Can use the Poisson distribution because the data is a count of a discrete random variable. We are assuming the events/counts are independent from one another. The events are being counted during a well defined interval.

- the mean (expected value) for the poisson distribution is equivalent to the lambda p parameter as well as the variance. 

## Exercises

1. What is the probability that she will hear more than 8 calls during any given session?

- Asking for the cumulative probability means we should call on the p-function associated with the poisson distribution as this will provide cumulative distribution probability values. 
  - This will be based on probability mass values as opposed to probability density values because we are working with discrete (not continuous) random variables. 

- The ppois will output the probability (area under the curve) up to (left of) the specified quantile. Since we know probabilities must sum to 1 we can subtract ppois output from 1 to get the area under the curve to the right of the count. 

##NK: It totally works to subtract ppois from 1. However, another option is to write lower = FALSE in the argument. This means rather than taking the area to the left, it will take the area to the right of the specified point. 

```{r}
1-ppois(q=8, lambda=15) #specifying q as 8 because we want MORE than (not including) 8. If it was more than or equal to 8 we would argue q as the realization one step down (aka 7)
```

What is the probability that she will hear no calls in a session?

- asking for probability of a singular realization so must call on the d-function which will output probability density values. 

```{r}
dpois(x=0, lambda=15) #if we rounded it to significant decimal places it would be 0% chance. 
```
##NK: for this one yours makes more sense. I used ppois which accomplished the same thing because x=0. However, for any other number I would've wanted to use dpois. 

What is the probability that she will hear exactly 3 calls in a session?

```{r}
dpois(3,15)
```
##NK: I did the same thing here. 

Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 30.

- This looks good as the poisson distribution probability mass values characteristically peak near the expected mean value and go to zero as x approaches infinity. 

```{r}
barplot(height=dpois(0:30,15), space=0, names.arg = 0:30, ylab = "Pr(X=x)", xlab='x', main= "Probability Mass Function of Titis Monkey Call Counts") 
# height = a vector to describe the bars of the plot. We use the dpois to calculate a vector of probability mass values for the counts 0 through 30 inclusive. 
#ylim is set from 0 to 1 because a probability must be within that range 
#names.arg is specifying the individual bar labels. In this case our count realizations (a sequence from 0-30)
```
#If you simply remove ylim, r automatically fits your data, so you have a lot less blank space. When you specify 1, it's hard to read your bar plot, but I'm not sure how to specifiy a value below 1. 

Simulate 104 results from this distribution (i.e., 2 years of Saturday monitoring sessions)

- random generation/simulation indicates we use the associated r-function

```{r}
set.seed(1) 
#use set seed to retain the exact random generation. 
rpois(n=104,15) 
#n argument for the rpois is indicating the number of variates you want to generate. #second argument is still the lambda value. 
```
##NK: Does it matter in this case whether or not you use set.seed? 

Plot the simulated results using hist() and use xlim() to set the horizontal limits to be from 0 to 30. How does your histogram compare to the shape of the probability mass function you plotted above?

- The histogram looks very similar to the previously plotted probability mass function indicating the generated data is indeed following a poisson distribution. Some differences are there is not a perfect peak at the lambda p parameter (15) and the plot is no longer perfectly symmetrical, but that is to be expected when working with actual data points over probabilities as probabilities are just an estimate of chance. 

```{r}
set.seed(1)
#calling set.seed to generate the same data to plot as was generated in the previous exercise
x <- rpois(n=104,15) 
#assigning the data to a random variable x to plug into the hist data argument
hist(x,xlim= c(0,30), main = "Histogram of Titus Monkey Call Counts", xlab= 'Count', ylab = 'Count Frequency')
```

## Challenges 

1. When using the histogram function I originally thought xlim needed to be denoted by the entire sequence 0:30. After looking on stack overflow and at the documentation the xlim argument needs to be a vector containing only the minimum and maximum value of a range. 
#I had this same issue! I also found the answer online somewhere. 
2. NA: After reading The Book of R Chapter 16 section on poisson distribution the exercises were extremely straight forward. (Even though I do have some questions still as to how the lambda value is generated or if that is something that will always be given in this class?)
#Your work was very easy to read. Overall, I think we approached it similarly. I think in the future, I will also copy in the question prompts, so people don't have to refer to the module when making comments. 